---
title: "mars()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Multivariate Adaptive Regression Spline (MARS) Models

Multivariatate Adaptive Regression Spline is a form of regression analysis introduced by Jerome H. Friedman in 1991. The model uses a hinge function as the interactions between bases functions, allowing a multi-dimensional model. 


# Getting Started
* Load the mars package. 
* Install the package via GitHub: https://github.com/SFUStatgen/SFUStat360Projects/tree/main/Projects/ProjectSY


```{r setup}
library(mars)
```


# Create mars object
The mars algorithm needs a control object called mars.control. 
Initialize the object first then pass the object into mars() function. 
```{r}
mc <- mars.control(Mmax = 8)

```

# Fit a mars model. 
For this example, we will use the mtcars data set from R.
We will use mpg (miles per gallon) as the response variable and try to fit a model using all other variables as explanatory variables. 
```{r}
mod.fit <- mars(mpg~., data = mtcars, control = mc)

```

We can also specify which explanatory variables to use:
```{r}
mars(mpg~ cyl+hp+drat+wt, data = mtcars, control = mc)



```


# Print the mars object
Once we fitted the model to the mtcars data, use `print()` method to print the model. 
```{r}
print(mod.fit)

```

The output of the print(mars) above is similar to that of the `print.lm()`.


# Summary 
We can print a summary of the fitted model using `summary()` method. `summary()` method takes an mars object as the argument. 

```{r}
summary(mod.fit)

```
The output from the `summary()` is very similar to that of the `summary.lm()`.


# Plot the mars object
The `plot()` method in the mars package allows us to visualize the multivariate model graphically. The arguement to the `plot()` is a mars object. 

```{r}

plot(mod.fit)

```

# Predict the mars object

Just like in the `lm()`, we can call the `predict()` to predict the response variable using the model. 
```{r}

#predict(mod.fit, newdata = mtcars)

```

# Reference

Jerome H. Friedman, The Annals of Statistics, Mar., 1991, Vol.19, No.1 (Mar., 1991), pp. 1-67